### 2025-04-05 causal-inference.qmd
20250405 - edits

In each of these examples, you are controlling for player education for purposes of examining the effect of sports drink consumption on performance.
Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox.
In the example, if we did not control for gender, we would infer that there is a positive association between dosage and recovery probability.
However, when we examine each men and women separately, we learn that the association between dosage and recovery probability is actually negative within each gender group.
Thus, in this case, failure to control for gender would lead to false inferences about the association between dosage and recovery probability.

>>>>

In each of these examples, you are controlling for player education for purposes of examining the effect of sports drink consumption on performance.
Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox, which is an example of [Simpson's paradox](#sec-simpsonsParadox) described in @sec-simpsonsParadox.
In the example, if we did not control for gender, we would infer that there is a positive association between dosage and recovery probability.
However, when we examine each men and women separately, we learn that the association between dosage and recovery probability is actually *negative* within each gender group.
Thus, in this case, failure to control for gender would lead to false inferences about the association between dosage and recovery probability.

### 2025-04-05 cognitive-bias.qmd
20250318 - increasing shot difficulty
20250405 - edits

@@

Evidence on the hot hand is mixed.
Considerable evidence historically has suggested that there is no such thing as a "hot hand" even though many players and fans believe in it [@Gilovich1985; @BarEli2006; @Moskowitz2011; @Avugos2013; @Wetzels2016].
Even though streaks may occur, making a shot may not increase the likelihood of making the next shot; that is, the likelihood of making a field goal seems to be independent of past success (or failure) on recent attempts.
Defenses may adjust to guard a "hot" player more closely.
However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a; @Miller2024].
However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].

>>>>

By contrast, according to @Lilienfeld2015a, some of the frequent features of pseudoscience include:

> 1. An overuse of ad hoc hypotheses designed to immunize claims from falsification
> 1. Absence of self-correction
> 1. Evasion of peer review
> 1. Emphasis on confirmation rather than refutation
> 1. Reversed burden of proof
> 1. Absence of connectivity
> 1. Overreliance on testimonial and anecdotal evidence
> 1. Use of obscurantist language
> 1. Absence of boundary conditions
> 1. The mantra of holism

@@

Evidence on the hot hand is mixed.
Considerable evidence historically has suggested that there is no such thing as a "hot hand" even though many players and fans believe in it [@Gilovich1985; @BarEli2006; @Moskowitz2011; @Avugos2013; @Wetzels2016].
Even though streaks may occur, making a shot may not increase the likelihood of making the next shot; that is, the likelihood of making a field goal seems to be independent of past success (or failure) on recent attempts.
Defenses may adjust to guard a "hot" player more closely.
However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a; @Miller2024].
However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].
Indeed, the slight potential increase in shooting ability tends to be offset by the tendency that "hot" players have to increase their shot difficulty after making a previous shot, so the net hot hand effect appears to be "vanishingly small" [@Partnow2021].

### 2025-04-05 correlation.qmd
20250405 - edits

The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.

@@

A nonlinear correlation can be estimated using the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:

>>>>

The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
As noted in @sec-statisticalTestsPartitionedVariance, the correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance, as in @eq-correlationPartitionedVariance.

@@

A nonlinear correlation can be estimated using the $\xi$ coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:

### 2025-04-05 multiple-regression.qmd
20250227 - clean up
20250405 - edits

@@

The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
Various formulas for $R^2$ are in @eq-rSquared.
Larger $R^2$ values indicate greater accuracy.
Multiple regression can be conceptualized with overlapping circles (similar to a venn diagram), where the non-overlapping portions of the circles reflect nonshared variance and the overlapping portions of the circles reflect shared variance, as in @fig-regression.

@@

Consider an example where you develop an algorithm to predict players' fantasy performance based on 2023 data using hundreds of [predictor variables](#sec-correlationalStudy).
To some extent, these [predictor variables](#sec-correlationalStudy) will likely account for true variance (i.e., signal) and error variance (i.e., noise).
If we were to apply the same algorithm based on the 2023 prediction model to 2024 data, the prediction model would likely predict less accurately than with 2023 data.
The regression coefficients in the

>>>>

### Evaluating the Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}

@@

As noted in @sec-statisticalTestsPartitionedVariance, 
The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions, as in @eq-coefficientOfDeterminationPartitionedVariance: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
Various formulas for $R^2$ are in @eq-rSquared.
Larger $R^2$ values indicate greater accuracy.
Multiple regression can be conceptualized with overlapping circles (similar to a venn diagram), where the non-overlapping portions of the circles reflect nonshared variance and the overlapping portions of the circles reflect shared variance, as in @fig-regression.

@@

Consider an example where you develop an algorithm to predict players' fantasy performance based on 2023 data using hundreds of [predictor variables](#sec-correlationalStudy).
To some extent, these [predictor variables](#sec-correlationalStudy) will likely account for true variance (i.e., signal) and error variance (i.e., noise).
If we were to apply the same algorithm based on the 2023 prediction model to 2024 data, the prediction model would likely predict less accurately than with 2023 data.
The regression coefficients in the [FILL IN]
